---
title: Annual income Variability between male and female UK Data Scientists and Job
  Satisfaction
author: "Safa A Elrais"
date: "10/05/2021"
output:
  pdf_document: default
  html_document:
    highlight: kate
    theme: cosmo
    css: mycss.css
    df_print: tibble
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(dplyr)
df <- read.csv("C:/Users/44751/OneDrive/Desktop/200266165/PSYC6422/2020.csv")
df <- df %>% 
  filter(Country == "United Kingdom")



```
## Introduction  :

### *Background *
Today, Data Science is one of the hottest Jobs in Technology with Data Scientists are among the highest paid with an average annual income of `r format (mean(df$ConvertedComp, na.rm = TRUE),digits = 5 )`£ in the UK. Sometimes the long working hours, the complexity of data problems, the high expectations from employees exert tremendous pressure on Data Scientists. The stressful nature of the job may affect the level of Job Satisfaction leading to many more Data Scientists leaving their jobs.    

###  *Research Questions*

By this Data Visualization, we intend to answer these questions :


1. How much satisfied are UK Data Scientists with their jobs ?
2. Is the high annual income justify the stressful nature of the job leading to more Job Satisfaction?
3. Is the high annual income more accounted for males or females Job Satisfaction in Data Science?

### *Data Origins*

The data is sourced directly from [Stack Overflow Annual Developer Survey](https://insights.stackoverflow.com/survey/) for the year 2020. The Survey consists of 6 sections and 61 variables . Respondents were from all over the world 


```{r}

#load the data

datafile= "C:/Users/44751/OneDrive/Desktop/200266165/PSYC6422/2020.csv"

df0 <- read.csv(datafile)

#Display the data

head(df0)


```
The analysis focuses mainly on 4 variables , the following table provides each variable names and its explanation.

```{r echo=FALSE}

 table <- tibble::tribble(
   ~Variable.name,                               ~Explanation,
         "Gender", "Either Man , Woman , Non-Binary",             
        "Country", "Country where respondent Currently lives",             
  "ConvertedComp", "Income per year",             
  "JobSat"       , "Level of satisfaction : either very satisfied,slightly satisfied , neither Satisfied nor dissatisfied , slightly disatisfied , very dissatisfied " ,   
  )

require(knitr)
kable(table, digits = 2, row.names = FALSE, align = "l",
              caption = NULL)


```
## Data cleaning and Preparation : 

The current data set has several issues before its ready for visualization , the following steps are required :

```{r echo=TRUE}

# Load libraries 

library(tidyverse)
library(ggplot2)
library(dplyr)

# Filter Data set to include United Kingdom observations only 

df0 <- df %>% 
  filter(Country== "United Kingdom")

# Select only the variables used in analysis

df <- df %>% 
  select ("Gender", "Country", "ConvertedComp", "JobSat")


# Remove the missing values from data set 

df <- na.omit(df)

# Subset Data TO include only Man and Woman respondents 

df <- subset(df, Gender == "Man" | Gender == "Woman")


 
# Show first 5 rows of the processed data 
 
head(df , 5)

# Summarize the processed data 

summary(df)


```


<font color = "Red"> *N.B : In my analysis I excluded 30 Observations from Gender Column to limit analysis only for respondents who identified themselves as Man or Woman .</font>


##Data Visualization :

##* Visualization 1 :  How much satisfied are UK Data Scientists with their jobs ? *

This Visualization shows how much satisfied are UK data Scientists in their careers 

```{r}

# label parameters of the bar graph

ylab <- "Percentage of Responses"
title <- "Overall  Career Satisfaction among UK Data Scientists "


#set the colors of the bar graph 


pallette <- c("Very satisfied" ="red",
              "Slightly satisfied"="blue",
               "Neither satisfied nor dissatisfied"="green",
               "Slightly dissatisfied"="yellow",
               "Very dissatisfied"= "grey")


# create the bar chart

library(ggplot2)

p <- ggplot(data=df, mapping = aes(x=JobSat, fill=("JobSat")))

p+geom_bar(mapping = aes(y=..prop..,group="JobSat"))+
  scale_x_discrete(limits=c("Very satisfied",
                            "Slightly satisfied",
                            "Neither satisfied nor dissatisfied",
                            "Slightly dissatisfied","Very dissatisfied"))+
  scale_y_continuous(labels = scales::percent) +labs(y=ylab,title = title)+
  coord_flip()+theme(legend.position = "none",panel.background = element_blank(),plot.title=element_text(family="Times", face="italic", size=18 , color = "#170630"), axis.title.y = element_blank(),axis.title.x = element_text(family= "Times", size = 10))+scale_fill_brewer(palette = "Dark2")
                          
                                                                                                                                                                                                                                
```

The above graph shows that majority of UK Data Scientists are satisfied with their jobs with around 65% of responses were either slightly or very satisfied. In the next visualization we will explore if this high level of satisfaction is related to high annual income or not .


##* Visualization 2 :Is the high annual income justify the stressful nature of the job leading to more Job Satisfaction? *


```{r}
library(ggplot2)

p2 <- ggplot(data=df , mapping =aes(x=JobSat , y=ConvertedComp , fill = JobSat))

p2+geom_violin() 

```

